Just Add IFC! A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/formal.tex
Pass
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

In this section we present the formal security guarantees provided by
our embedding.
%
Concretely, we show that |specLang alpha targetLang| for, appropriate
scheduler, satisfies non-interference~\cite{Goguen82},
without making any reference to properties of |targetLang|.
%
%As it turns out, 
The precise scheduling policy |alpha| dictates
the guarantee holding for programs with diverging tasks.
For a sequential scheduler |seqf|, we will only be able to show
\emph{termination-insensitive non-interference}, where a program
may diverge based on secret data; this allows attackers to observe
secrets by observing the termination of tasks.
For the concurrent round-robin schedule |roundrobinf|,
we can show a stronger result known as
\emph{termination-sensitive non-interference},
where termination attacks cannot leak information.

How can we characterize what the \emph{secret inputs} of a program are?  Like
other work~\cite{Li+:2010:arrows,Russo+:Haskell08,lio,stefan:addressing-covert},
we specify and prove non-interference using \emph{term erasure}.
%
%Intuitively, we prove that an attacker does not learn
%any sensitive information from a program by 
Intuitively, term erasure allows us to show that a an attacker does not learn
any sensitive information from a program if the program behaves identically
(from the attacker point of view) to a program with all sensitive data
``erased''.
%
To accommodate for this, we extend our formalism to account for
programs with erased configurations and a function |erasef il| that
performs erasures, usually by rewriting (parts of) configurations that
are more sensitive than |il| to a new syntactic construct |bullet|.
%
We can characterize the types of languages for which one can state
a non-interference claim.

\Red{should be the environment colored in the definition?}
\begin{definition}[Information flow control language]
    An information flow control language |L| is a tuple |(C, .->,
    erasef l)|, where $|C|$ is the type of machine configurations (members
    of which are usually denoted by the meta-variable |c|), |.->| is a
    reduction relation between machine configurations and |erasef l : C -> tyerase C|
    is an erasure function parametrized on labels from machine configurations to \emph{erased} machine
    configurations |tyerase C|.  This tuple is sometimes
    augmented with a set of terminal configuration states $V
    \subseteq |C|$ for which there are no possible further transitions.
    \Red{It is a bit odd to extend a tuple ``sometimes'' and a |V| appears like from
      nowhere. What about saying: Sometimes, we refer as |V| to set of 
     terminal configurations in |C|, i.e., configurations states where
     no further transitions are possible. If you agree, then the next sentence
     where you said |V = ...| must be removed.}
\end{definition}

%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
Our language |specLang alpha (targetLang nop)| can be
represented as |(ifc C, ifc V, .->, erasef il)|, such that
$|ifc C| = |iS| \times \prod |it|$ and
$|ifc V| = |iS| \times |it|_V$, where $|it|_V \subset |it|$ is the
type for tasks whose expressions have been reduced to
values.\footnote{
  Here, we abuse notation by describing types for configuration parts using the
  same metavariables as the ``instance'' of the type, e.g., |it| for the type of
  task.
}
The erased configuration |tyerase (ifc C)| extends |ifc C| with configurations
containing |bullet|, and Figure~\ref{fig:erasure} gives the precise definition for
our erasure function |erasef il|.
%
A task and its corresponding message queue is completely removed from the task
list if its label does not flow to the attacker observation level |il|.
Otherwise, we apply the erasure function homomorphically and remove any message
from the task's message queue that are more sensitive than |il|.
%}

@@

We note that the definition of |erasef il| captures the IFC
language's attacker model. 
%
In our case, we assume that the attacker cannot observe sensitive tasks or
messages, or even just the number of such entities.
%
While such assumptions are standard~\tocite{}, our definitions allow for
stronger attackers that may be able to inspect resource usage.\footnote{
  We believe that we can extend |specLang alpha (targetLang nop)| to
  such models using the resource limits techniques of~\tocite{}.
  %
  We leave this extension to future work.
}

@@

\begin{definition}[Termination Sensitive Non-Interference (TSNI)]
  A language |(C, .->, erasef l)| is termination
  sensitive non-interfering if for any label |l|, and configurations
  |c1, c1', c2 member C|, if
  \begin{equation} \label{eq:tsni-lhs}
    |c1| \approx_{|l|} |c2|
    \qquad \text{and} \qquad
    |c1| |.->|^* |c1'|
  \end{equation}
  then there exists a configuration |c2' member C| such that
  \begin{equation} \label{eq:tsni-rhs}
    |c1'| \approx_{|l|} |c2'|
     \qquad \text{and} \qquad
    |c2| |.->|^* |c2'|
    \ \text{.}
  \end{equation}
\end{definition}
%
In other words if we take two |l|-equivalent configurations, then for every
intermediate step taken by the first configuration, there is a corresponding
number of steps that the second configuration can take to result in a
configuration that is |l|-equivalent to the first resultant configuration.
\Red{TODO: Do we want a commutative diagram to illustrate this? Ale: I don't
  think so, it is quite standard}

>>>>

In this section, we present the formal security guarantees provided by
our embedding.
%
Concretely, we show that |specLang alpha targetLang|, with an appropriate
scheduler |alpha|, satisfies non-interference~\cite{Goguen82},
without making any reference to properties of |targetLang|.
%
The precise scheduling policy dictates
what guarantee we can achieve for programs with diverging tasks.
For a sequential scheduler |seqf|, we will only be able to show
\emph{termination-insensitive non-interference}, where a program
may diverge based on secret data; this allows attackers to observe
secrets by observing the termination of tasks.
For the concurrent round-robin schedule |roundrobinf|,
we can show a stronger result known as
\emph{termination-sensitive non-interference},
where termination attacks cannot leak information.

How can we characterize what the \emph{secret inputs} of a program are?  Like
other work~\cite{Li+:2010:arrows,Russo+:Haskell08,lio,stefan:addressing-covert},
we specify and prove non-interference using \emph{term erasure}.
%
Intuitively, term erasure allows us to show that an attacker does not learn
any sensitive information from a program if the program behaves identically
(from the attacker point of view) to a program with all sensitive data
``erased''.
%
To accommodate for this, we extend our formalism with a function |erasef l| that
performs erasures by mapping configurations to erased configurations,
usually by rewriting (parts of) configurations that are more sensitive
than |l| to a new syntactic construct |bullet|.
%
In order for us to state non-interference, a language needs to include an
erasure function:

%\Red{should be the environment colored in the definition?}
%EZY: No, I don't think so.
\begin{definition}[Information flow control language]
    An information flow control language |L| is a tuple |(C, .->,
    erasef l)|, where $|C|$ is the type of machine configurations (members
    of which are usually denoted by the meta-variable |c|), |.->| is a
    reduction relation between machine configurations and |erasef l : C -> tyerase C|
    is an erasure function parametrized on labels from machine configurations to \emph{erased} machine
    configurations |tyerase C|.  Sometimes, we refer as |V| to set of
     terminal configurations in |C|, i.e., configurations states where
     no further transitions are possible.
\end{definition}

%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
Our language |specLang alpha (targetLang nop)| can be
represented as |(ifc C, .->, erasef il)|, such that
$|ifc C| = |iS| \times \prod |it|$.  Here,
$|ifc V| = |iS| \times |it|_V$, where $|it|_V \subset |it|$ is the
type for tasks whose expressions have been reduced to
values.\footnote{
  Here, we abuse notation by describing types for configuration parts using the
  same metavariables as the ``instance'' of the type, e.g., |it| for the type of
  task.
}
The erased configuration |tyerase (ifc C)| extends |ifc C| with configurations
containing |bullet|, and Figure~\ref{fig:erasure} gives the precise definition for
our erasure function |erasef il|.
%
A task and its corresponding message queue is completely removed from the task
list if its label does not flow to the attacker observation level |il|.
Otherwise, we apply the erasure function homomorphically and remove any message
from the task's message queue that are more sensitive than |il|.
%}

@@

The definition of |erasef il| is quite important, as it captures the IFC
language's attacker model: elements that are erased cannot be observed
by the attacker.
%
In our case, we assume that the attacker cannot observe sensitive tasks or
messages, or even the number of such entities.
%
While such assumptions are standard~\tocite{}, our definitions allow for
stronger attackers that may be able to inspect resource usage.\footnote{
  We believe that we can extend |specLang alpha (targetLang nop)| to
  such models using the resource limits techniques of~\tocite{}.
  %
  We leave this extension to future work.
}

@@

\begin{definition}[Termination Sensitive Non-Interference (TSNI)]
  A language |(C, .->, erasef l)| is termination
  sensitive non-interfering if for any label |l|, and configurations
  |c1, c1', c2 member C|, if
  \begin{equation} \label{eq:tsni-lhs}
    |c1| \approx_{|l|} |c2|
    \qquad \text{and} \qquad
    |c1| |.->|^* |c1'|
  \end{equation}
  then there exists a configuration |c2' member C| such that
  \begin{equation} \label{eq:tsni-rhs}
    |c1'| \approx_{|l|} |c2'|
     \qquad \text{and} \qquad
    |c2| |.->|^* |c2'|
    \ \text{.}
  \end{equation}
\end{definition}
%
In other words if we take two |l|-equivalent configurations, then for every
intermediate step taken by the first configuration, there is a corresponding
number of steps that the second configuration can take to result in a
configuration that is |l|-equivalent to the first resultant configuration.
